---
output:
  pdf_document: default
  html_document: default
---
---
title: "Assignment 2"
author: "Arvid Gramer and Anton Linnér"
date: "`r Sys.Date()`"
output: pdf_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
set.seed(2) 
library(copula)
library(scatterplot3d)
library(ggplot2)
library(reshape2)
```
\newpage
\section{Exercise 1}
\textbf{1:}
An Archimedean copula is a copula on the form: $C(u_1,..., u_n) =       \varphi(\sum^n_{i=1}\varphi^{-1}(u_i))$, where $\varphi$ is the copula's generator function. The conditions for this generator function is that \begin{itemize}
        \item $\varphi(0) = 1,$ $\varphi(\infty) = 0$
        \item $\varphi$ is decreasing
        \item $\varphi$ is continuous
\end{itemize} 
A convenient way of obtaining such generator functions is to take the Laplace transform of a random variable, $\varphi_X(t) = \int_0^{\infty}\mathrm{e}^{-tx}f_X(x)dx$. Depending on choice of generator function we retrieve different families of Archimedean copulas, such as Clayton, Gumbel or Frank copulas.

\textbf{2:}
Frank copula: This copulas has generator function $\varphi(t) = \mathrm{ln}\left(\frac{\mathrm{e}^{-\theta t}-1}{\mathrm{e}^{-\theta}-1}\right)$, C(u,v) = $\varphi^{-1}(\varphi(u) + \varphi(v)) = \frac{1}{\theta}\mathrm{ln}\left(1+\frac{(\mathrm{e}^{-\theta u}-1)(\mathrm{e}^{-\theta v}-1)}{\mathrm{e}^{-\theta}-1}\right)$

Clayton copula: The generator is $\varphi(t) = t^{-\theta}-1$, which results in a copula $C(u,v) = (u^{-\theta}+v^{-\theta}-1)^{-1/\theta}$
\section{Exercise 2}

\subsection{Exercise 2.1}
We can see that Spearmans Rho and normal correlation are somewhat similar, but Kendalls Tau is lower than both. We also see that the theoretical values are very close to the empirical values for all methods. And lastly we see that when we rank the observations, Spearmans rho gives the same value as when we didn't rank the values. 
```{r Exercise 2.1, echo = TRUE, include = TRUE}
myCop.norm <- ellipCopula(family="normal", dim=3, dispstr="ex", param = 0.4)
x1 <- rCopula(500, myCop.norm)
scatterplot3d(x1)
#Correlation
(rho1 = cor(x1[,1], x1[,2]))
(rho2 = cor(x1[,1], x1[,3]))
(rho3 = cor(x1[,2], x1[,3]))

#Kendall
(cor(x1[,1], x1[,2], method="kendall")) 
(cor(x1[,1], x1[,3], method="kendall"))
(cor(x1[,2], x1[,3], method="kendall"))
#Theoretical
(t1_theo = 2/pi*asin(0.4)) 

#Spearman
(cor(x1[,1], x1[,2], method="spearman"))
(cor(x1[,1], x1[,3], method="spearman"))
(cor(x1[,2], x1[,3], method="spearman"))
#Theoretical
rho_s_theo = 6/pi*asin(0.2)

#Rho 
(rho_Ss1 = cor(rank(x1[,1]), rank(x1[,2])))
(rho_Ss2 = cor(rank(x1[,1]), rank(x1[,3])))
(rho_Ss3 = cor(rank(x1[,2]), rank(x1[,3])))

```
\newpage

\subsection{Exercise 2.2}
```{r Exercise 2.2a,echo = TRUE, include = TRUE}
myCop.t1 <- ellipCopula(family="t", dim=2, dispstr="ex", param = -0.4, df = 8)
myCop.t2 <- ellipCopula(family="t", dim=2, dispstr="ex", param = 0, df = 8)
myCop.t3 <- ellipCopula(family="t", dim=2, dispstr="ex", param = 0.4, df = 8)

x2 = rCopula(500, myCop.t1)
x3 = rCopula(500, myCop.t2)
x4 = rCopula(500, myCop.t3)
```


As in previously when Gaussian copula was used, the methods seem to give results that are in accordance with the theoretical results. 
``` {r Exercise 2.2b, echo = TRUE, include = TRUE}
plot(x2)
plot(x3)
plot(x4)

(rho1 = cor(x2[,1], x2[,2]))
(rho2 = cor(x3[,1], x3[,2]))
(rho3 = cor(x4[,1], x4[,2]))

#Kendall

#Empirical
(cor(x2[,1], x2[,2], method="kendall")) 
(cor(x3[,1], x3[,2], method="kendall"))
(cor(x4[,1], x4[,2], method="kendall"))
#Theoretical for the three cases
(t1_theo = 2/pi*asin(-0.4)) 
(t1_theo = 2/pi*asin(0)) 
(t1_theo = 2/pi*asin(0.4)) 

#Spearman
#Empirical
(cor(x2[,1], x2[,2], method="spearman")) 
(cor(x3[,1], x3[,2], method="spearman"))
(cor(x4[,1], x4[,2], method="spearman"))
#Theoretical
(rho_s_theo = 6/pi*asin(-0.2))
(rho_s_theo = 6/pi*asin(0))
(rho_s_theo = 6/pi*asin(0.2))

```
\subsection{Exercise 2.3}
By looking at the appendix and solving for the expression for the $\tau$ using $\theta$ = 3 and k = 1. Using this, we get a value for $\tau$ thats about 0.3, which is in agreement with what is calculated below. 
```{r Exercise 2.3,echo = TRUE, include = TRUE}
frank.cop <- frankCopula(dim = 3, param=3)
x5 = rCopula(500, frank.cop)

(rho1_x5 = cor(x5[,1], x5[,2]))
(rho2_x5 = cor(x5[,1], x5[,3]))
(rho3_x5 = cor(x5[,2], x5[,3]))

(t1_x5 = cor(x5[,1], x5[,2], method = "kendall"))
(t2_x5 = cor(x5[,1], x5[,3], method = "kendall"))
(t3_x5 = cor(x5[,2], x5[,3], method = "kendall"))

(rho_s1 = cor(x5[,1], x5[,2], method = "kendall"))
(rho_s2 = cor(x5[,1], x5[,3], method = "kendall"))
(rho_s3 =cor(x5[,2], x5[,3], method = "kendall"))

```
\subsection{Exercise 2.4}
Below is the code for the function:
```{r Exercise 2.4, echo = TRUE, include = TRUE}
gammaRand <- function(n,theta){
   B= 1/theta
   M=runif(n)
   E = rexp(2*n)
   dim(E) = c(2,n)
   E=t(E)
   U = E/M
   sample = B/(U+B)
  
}
gammaRand(500, 0.5)
gammaRand(500, 1)
gammaRand(500, 5)
gammaRand(500, 50)

```
\subsection{Exercise 2.5}
The reason T(x) is increasing is due to the fact that arctanh is increasing in this interval (between -1 and 1). 
We get -1 and 1 as dependace for Kendalls Tau and Spearman, but only -0.914 and + 0.914 for normal correlation. Since Z and Y are created directly from X, it is reasonable that they should have perfect dependance. How ever, since there is a non-linear transform, normal Rho wont be quite as good when trying to measure this depndance. 

```{r Exercise 2.5,echo = TRUE, include = TRUE}
X = runif(1000)
Y= atanh(X)
Z=-Y
par(mfrow=c(2,1))
plot(X,Y)
plot(X,Z)
cor(X,Y, method="kendall")
cor(X,Z, method = "kendall")

cor(X,Y, method= "spearman")
cor(X,Z, method = "spearman")

cor(X,Y)
cor(X,Z)

```
\section{Exercise 3}
The plots for the t-copula. For the first plots (scatter, cdf, pdf) we can see that the plots show more dependance, which is reasonable since this is how they were created. 
```{r Exercise 3,echo = TRUE, include = TRUE}
par(mfrow= c(1,3)) #mfrow = c(1,2) is one row two cols.


plot(x3)
persp(myCop.t1, pCopula)
contour(myCop.t1, dCopula)


plot(x4)
persp(myCop.t2, pCopula)
contour(myCop.t2, dCopula)
```
\section{Exercise 4}
\subsection{Scatterplots}
The plots are, in order, Gaussian, Gumble, Clayton and t-copula. We can see that there is a somewhat similar dependance in all of the plots. We can see in plot 3 (Clayton), it seems to have quite a strong lower-tail dependance. The t-copula seems to have stronger tail dependance than the Gaussian, which might be expected. The Gumble copula seems to have slightly stronger upper tail dependance.
```{r Exercise 4,echo = TRUE, include = TRUE}
# Copulas
gauss.cop <- ellipCopula(family="normal", dim=2, dispstr="ex", param = 0.7) 
gumbel.cop <- archmCopula("gumbel",2 )
clayton.cop <- claytonCopula(2.2, dim = 2)
myCop.t <- ellipCopula(family="t", dim=2, dispstr="ex", param = 0.71, df = 4)
#Samples
x6= rCopula(2000, gauss.cop)
x7= rCopula(2000, gumbel.cop)
x8= rCopula(2000, clayton.cop)
x9= rCopula(2000, myCop.t)

#Plots
par(mfrow= c(2,2), mar = c(2,2,1,1), oma=c(1,1,0,0), mgp = c(2,1,0))
plot(x6)
plot(x7)
plot(x8)
plot(x9)

```
\newpage
\subsection{Quantile transformation}
What we can see in these plots is that we still have the same behaviour which was described in the previous section, only slightly clearer. 
```{r Quantile transformation,echo = TRUE, include = TRUE}
par(mfrow= c(2,2), mar = c(2,2,1,1), oma=c(1,1,0,0), mgp = c(2,1,0))
x6= rCopula(2000, gauss.cop)
x7= rCopula(2000, gumbel.cop)
x8= rCopula(2000, clayton.cop)
x9= rCopula(2000, myCop.t)

x6[,1] = rank(x6[,1])/length(x6[,1])
x6[,2] = rank(x6[,2])/length(x6[,2])

x6[,1] <- qnorm(x6[,1])
x6[,2] <- qnorm(x6[,2])



x7[,1] = rank(x7[,1])/length(x7[,1])
x7[,2] = rank(x7[,2])/length(x7[,2])

x7[,1] <- qnorm(x7[,1])
x7[,2] <- qnorm(x7[,2])


x8[,1] = rank(x8[,1])/length(x8[,1])
x8[,2] = rank(x8[,2])/length(x8[,2])

x8[,1] <- qnorm(x8[,1])
x8[,2] <- qnorm(x8[,2])

x9[,1] = rank(x9[,1])/length(x9[,1])
x9[,2] = rank(x9[,2])/length(x9[,2])

x9[,1] <- qnorm(x9[,1])
x9[,2] <- qnorm(x9[,2])
plot(x6)
plot(x7)
plot(x8)
plot(x9)

```
\section{Exercise 5}

As we can see, when we use the Gumbel copula with either two exponential margins or two standard normal margins, we get a similar Spearman Rho (around 0.65 to 0.75). 

```{r Exercise 5,echo = TRUE, include = TRUE}
cop_dist1 <- mvdc(copula=ellipCopula(family="norm", dim = 2, param=0.75), margins=c("norm", "exp"), paramMargins = list(list(mean=0,sd=2), list(rate=2))) 
cop_dist2 <- mvdc(copula=archmCopula("gumbel",dim=2, 2), margins=c("exp", "exp"), paramMargins = list(list(rate=4), list(rate=2)))

x10 = rMvdc(500, cop_dist1)
x11 = rMvdc(500, cop_dist2)
# First copula
cor(x10[,1],x10[,2], method = "spearman")
cor(x10[,1],x10[,2])

#Second copula
cor(x11[,1],x11[,2])
cor(x11[,1],x11[,2], method = "spearman")

#New margins to compare with second copula
cop_dist3 <- mvdc(copula=archmCopula("gumbel",dim=2, 2), margins=c("norm", "norm"), paramMargins = list(list(mean=0,sd=1), list(mean=0,sd=1)))
x12 = rMvdc(500, cop_dist3)
cor(x12[,1],x12[,2])
cor(x12[,1],x12[,2], method = "spearman")

```
\newpage
\section{Exercise 6}
\subsection*{Plots}
```{r exercise 6,echo = TRUE, include = TRUE}
par(mfrow= c(3,2))
plot(x10)
contour(cop_dist1, dMvdc, xlim=c(-6, 4), ylim=c(-0.1, 1.5))
plot(x11)
contour(cop_dist2, dMvdc, xlim = c(-0,0.35), ylim=c(-0,0.5))
plot(x12)
contour(cop_dist3, dMvdc, c(-2,2), ylim=c(-2,2))
```

\newpage
\section{Exercise 7}
We can see that the correlation of price and horsepower is high and positive, so it does seen to be the case that vehicles with higher horsepower are more costly. Simliarly, MPGcity and price have a negative correlation. 
```{r Exercise 7 car data,echo = TRUE, include = TRUE}
car_data <- read.csv("C:/Users/anton/Desktop/University/FMSN65  MASM33 - Quantitative Risk Management Using Copulas/Projects/Ass. 2/cardata.txt")

#Pairwise scatterplot of the data
pairs(car_data)
#Spearman
cor(car_data, method = "spearman")
cor.test(car_data[,1],car_data[,3], method = "spearman")
cor.test(car_data[,1],car_data[,2], method = "spearman")

```
\newpage
\section{Exercise 8}
We see that the probabilites of exceeding both thresholds is around 6-8 %, when using the 95th-percentile and 1-2 % for the 99th-percentile. We can also see that the ratios don't change when we change the threshold. We also see that Gumble copula gives the highest value for expected loss at 0.395, and the t-copula gives the lowest expected value at around 0.354. \newline
We can also see from the tail-dependance plots that the Gauss copula shows quite poor dependance. But the Gumble and t-copula shows some upper tail dependance. This makes it more reasonable to use these two copulas if what you are interested in is in the tails of a distribution. An example might be when dealing with stocks. 
```{r Exercise 8,echo = TRUE, include = TRUE}
Gauss <- mvdc( ellipCopula(family="normal", dim=2, dispstr="ex", param = 0.7), margins=c("lnorm", "lnorm"), paramMargins = list(list(mean=0, sd=1), list(mean=0, sd=1)))

t_cop = mvdc( ellipCopula(family="t", dim=2, dispstr="ex", param = 0.7, df = 2), margins=c("lnorm", "lnorm"), paramMargins = list(list(mean=0, sd=1), list(mean=0, sd=1)))

gumb = mvdc( archmCopula("gumbel" , 2), margins=c("lnorm", "lnorm"), paramMargins = list(list(mean=0, sd=1), list(mean=0, sd=1)))



#The threshould is:
(u = qlnorm(0.95,0,1))
(prob_gauss = 1- pMvdc(c(u,u), Gauss))
(prob_t = 1- pMvdc(c(u,u), t_cop))
(prob_gumb =1- pMvdc(c(u,u), gumb))

(r1 = prob_gauss/prob_t)
(r2 = prob_gauss/prob_gumb)
(r3 = prob_t/prob_gumb)
#The 99th percentile is:
(u = qlnorm(0.99,0,1))

#The probabilites of exceeding both thresholds:
(prob_gauss.a = 1- pMvdc(c(u,u), Gauss))
(prob_t.b = 1- pMvdc(c(u,u), t_cop))
(prob_gumb.c =1- pMvdc(c(u,u), gumb))

(r1a = prob_gauss/prob_t)
(r2b = prob_gauss/prob_gumb)
(r3c = prob_t/prob_gumb)

#t-copula
loss = rMvdc(10000, t_cop)
loss[which(loss[,1] <= u & loss[,2] <= u), ] <- 0
loss = rowSums(loss)
(E.t = mean(loss))

#Gauss
loss = rMvdc(10000, Gauss)
loss[which(loss[,1] <= u & loss[,2] <= u), ] <- 0
loss = rowSums(loss)
(E.Gauss = mean(loss))

#Gumble
loss = rMvdc(10000, gumb)
loss[which(loss[,1] <= u & loss[,2] <= u), ] <- 0
loss = rowSums(loss)
(E.gumb = mean(loss))


par(mfrow= c(1,3))
u = qlnorm(0.95,0,1)

loss = rMvdc(100000, Gauss)
tail.gauss = loss[apply(loss > u, 1, all), ]
plot(tail.gauss,ylim=c(0,100), xlim=c(0,100))

loss = rMvdc(100000, gumb)
tail.gumb = loss[apply(loss > u, 1, all), ]
plot(tail.gumb,ylim=c(0,100), xlim=c(0,100))

loss = rMvdc(100000, t_cop)
tail.t_cop = loss[apply(loss > u, 1, all), ]
plot(tail.t_cop,ylim=c(0,100), xlim=c(0,100))
```

```{r Exercise 9 FML}
library(fitdistrplus)
#Load data
ins.data <- read.csv("C:/Users/anton/Desktop/University/FMSN65  MASM33 - Quantitative Risk Management Using Copulas/Projects/Ass. 2/insurance.txt",sep="")
plot(ins.data[,1], ins.data[,2], log="xy")
cor9 = cor(ins.data[,1], ins.data[,2])

#First margin (of log(data))

par(mfrow= c(2,1))
hist(ins.data[,1], breaks=60)
margin1 = mledist(ins.data[,1], "lnorm")
margin1$estimate

test1 = rlnorm(length(ins.data[,1]),margin1$estimate[1], margin1$estimate[2])
hist(test1, breaks=100)

#Second margin:

hist(ins.data[,2], breaks=500)
margin2 = mledist(ins.data[,2], "lnorm")
margin2$estimate

test2 =rlnorm(length(ins.data[,1]),margin2$estimate[1], margin2$estimate[2])
hist(test2, breaks=500)

#Create/test a copulas:
cop1 =  mvdc(archmCopula("frank",2 ), margins=c("lnorm", "lnorm"), paramMargins=list(list(mean=margin1$estimate[1],sd=margin1$estimate[2]), list(mean=margin2$estimate[1],sd=margin2$estimate[2])))#FML

#Create starting point (b stands for margin params, a stands for copula params)
loglikMvdc(start, ins.data, cop1)
mm <- apply(log(ins.data), 2, mean)
vv <- apply(log(ins.data), 2, var)
(b1.0 <-c(mm[1],vv[1]))
(b2.0 <- c(mm[2],vv[2]))
(a.0 <- sin(cor(ins.data[, 1], ins.data[, 2], method = "kendall") * pi/2))
start <- c(b1.0, b2.0, a.0)
fit <- fitMvdc(ins.data, cop1,start=start, optim.control = list(trace = TRUE, maxit = 2000))

```

```{r Exercise 9 FML test 2}
# library(fitdistrplus)
# #Load data
# ins.data <- read.csv("C:/Users/anton/Desktop/University/FMSN65  MASM33 - Quantitative Risk Management Using Copulas/Projects/Ass. 2/insurance.txt",sep="")
# plot(ins.data[,1], ins.data[,2], log="xy")
# cor9 = cor(ins.data[,1], ins.data[,2])
# dat = log(ins.data)
# 
# #First margin (of log(data))
# 
# par(mfrow= c(2,1))
# hist(dat[,1], breaks=60)
# margin1 = mledist(dat[,1], "norm")
# margin1$estimate
# 
# test1 = rnorm(length(dat[,1]),margin1$estimate[1], margin1$estimate[2])
# hist(test1, breaks=100)
# 
# #Second margin:
# 
# hist(dat[,2], breaks=500)
# margin2 = mledist(dat[,2], "norm")
# margin2$estimate
# 
# test2 =rnorm(length(ins.data[,1]),margin2$estimate[1], margin2$estimate[2])
# hist(test2, breaks=500)
# 
# #Create/test a copulas:
# cop1 =  mvdc(archmCopula("gumbel",3 ), margins=c("norm", "norm"), paramMargins=list(list(mean=margin1$estimate[1],sd=margin1$estimate[2]), list(mean=margin2$estimate[1],sd=margin2$estimate[2])))
# #FML
# #Create starting point (b stands for margin params, a stands for copula params)
# 
# start2=c(margin1$estimate[1], margin1$estimate[2],margin2$estimate[1], margin2$estimate[2],0.5)
# loglikMvdc(start, dat, cop1)
# mm <- apply(dat, 2, mean)
# vv <- apply(dat, 2, var)
# (b1.0 <-c(mm[1],vv[1]))
# (b2.0 <- c(mm[2],vv[2]))
# (a.0 <- 1/(1-cor(dat[, 1], dat[, 2], method = "kendall")))
# start <- c(b1.0, b2.0, a.0)
# fit <- fitMvdc(dat, cop1, start = start, optim.control = list(trace = TRUE, maxit = 2000))

```

```{r Exercise 9 FML test 2}
library(fitdistrplus)
#Load data
par(mfrow=c(1,1))
dat <- read.csv("C:/Users/anton/Desktop/University/FMSN65  MASM33 - Quantitative Risk Management Using Copulas/Projects/Ass. 2/insurance.txt",sep="")
#dat = log(dat)
plot(dat[,1], dat[,2])
(cor9 = cor(ins.data[,1], ins.data[,2], method="kendall"))
# min_max_scale <- function(x) {
#   return((x - min(x)) / (max(x) - min(x)))
# }
# 
# # Apply min-max scaling to each column
# dat_scaled <- as.data.frame(lapply(dat, min_max_scale))
# plot(dat_scaled[,1], dat_scaled[,2], log="xy")

#First margin (of log(data))

par(mfrow= c(2,1))
hist(dat[,1], breaks=60)
margin1 = mledist(dat[,1], "lnorm")
margin1$estimate
test1 = rlnorm(length(dat[,1]),margin1$estimate[1], margin1$estimate[2])
hist(test1, breaks=100)

#Second margin:
hist(dat[,2], breaks=500)
margin2 = mledist(dat[,2], "lnorm")
margin2$estimate
test2 =rlnorm(length(ins.data[,1]),margin2$estimate[1], margin2$estimate[2])
hist(test2, breaks=100)




#Create/test a copulas:
cop1 =  mvdc(archmCopula("gumbel",5), margins=c("lnorm", "lnorm"), paramMargins=list(list(mean=0,sd=1), list(mean=0,sd=1)))#FML

cop2 =  mvdc(archmCopula("clayton",0.9 ), margins=c("lnorm", "lnorm"), paramMargins=list(list(mean=margin1$estimate[1],sd=margin1$estimate[2]), list(mean=margin2$estimate[1],sd=margin2$estimate[2])))


test3 = rMvdc(1500, cop1)
plot(test3, xlim = c(0, 2*10^6), ylim = c(0, 2*10^5))
plot(dat,  xlim = c(0, 2*10^6),ylim = c(0, 2*10^5))
#Create starting point (b stands for margin params, a stands for copula params)
start = c(margin1$estimate, margin2$estimate, 2)
loglikMvdc(start, as.matrix(dat), cop1)

fit <- fitMvdc(as.matrix(dat), cop1, start = start, optim.control = list(trace = TRUE, maxit = 2000))
fit2 = fitMvdc(as.matrix(dat), cop1, start = start, optim.control = list(trace = TRUE, maxit = 2000))
fit3 = fitMvdc(as.matrix(dat), cop1, start = start, optim.control = list(trace = TRUE, maxit = 2000))

test3 = rMvdc(1500, cop1)

cop3 =  mvdc(archmCopula("gumbel",1.468 ), margins=c("lnorm", "lnorm"), paramMargins=list(list(mean=9.373,sd= 1.670 ), list(mean=8.523 ,sd=1.428 )))
test3 = rMvdc(1500, cop3)
plot(test3, xlim = c(0, 2*10^6), ylim = c(0, 2*10^5))
plot(dat,  xlim = c(0, 2*10^6),ylim = c(0, 2*10^5))

```




```{r}
library(copula)
loglik.marg <- function(b, x) sum(dlnorm(x, mean = b[1], sd = b[2]))
ctrl <- list(fnscale = -1)
margin1Fit = mledist(dat[,1] , "lnorm")
b1 = margin1Fit$estimate

margin2Fit = mledist(dat[,2], "lnorm")
b2 = margin2Fit$estimate
udat <- cbind(plnorm(dat[, 1], mean = b1[1], sd = b1[2]),
 plnorm(dat[, 2], mean = b2[1], sd = b2[2]))
fit.ifl <- fitCopula( cop3@copula,udat, start =2)

#Params
params = c(b1,b2,fit.ifl@estimate)


```



```{r trying stuff}
dat <- read.csv("C:/Users/anton/Desktop/University/FMSN65  MASM33 - Quantitative Risk Management Using Copulas/Projects/Ass. 2/insurance.txt",sep="")
#Skapar dist där jag använder gumbel copula och några lnorm marginaler.
mvdc1 =  mvdc(archmCopula("gumbel",5), margins=c("lnorm", "lnorm"), paramMargins=list(list(mean=0,sd=1), list(mean=0,sd=1)))

# Hittar ML estimates för enbart margins, kommer användas vid någon tidspunkt. Tyckte "lnorm" passade marginsen rätt bra när man plottade hist av data. 
margin1Fit = mledist(dat[,1], "lnorm")
margin1Fit$estimate
margin2Fit = mledist(dat[,2], "lnorm")
margin2Fit$estimate

#Sätter några startvärden (det sista margins parametrar först, copulans paramter efter.)
start=c(9,1.5,8.5,1.4,5)
#FML fit. Alla parametrar av min dist fittas direkt. (går inte om jag använder marginfitsen ovan som startpunkt, av någon konstig anledning)
fit = fitMvdc(as.matrix(dat), mvdc1, start = start, optim.control=list(trace=TRUE, maxit = 2000))

#Testar även en Frank copula, med samma margins
mvdc2 =  mvdc(archmCopula("frank",5), margins=c("lnorm", "lnorm"), paramMargins=list(list(mean=0,sd=1), list(mean=0,sd=1)))
start=c(9,1.5,8.5,1.4,5)
fit = fitMvdc(as.matrix(dat), mvdc2, start = start, optim.control=list(trace=TRUE, maxit = 2000))
#Testar clayton copula
mvdc3 =  mvdc(archmCopula("clayton",5), margins=c("lnorm", "lnorm"), paramMargins=list(list(mean=0,sd=1), list(mean=0,sd=1)))
start=c(9,1.5,8.5,1.4,5)
fit = fitMvdc(as.matrix(dat), mvdc3, start = start, optim.control=list(trace=TRUE, maxit = 2000))


#Kör goodness of fit. Hur ska man tolka det. 
gof1 = gofCopula(fit@mvdc@copula, as.matrix(dat))
```






