---
title: "Assignment3"
author: "Anton Linnér"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(evd)
library(SimCop)
library(copula)
```

\\section(Random number generation from bivariate EVD) \\subsection(Parametric bivariate EV models) In the package \texttt{evd} the nine models and their respective dependence and asymmetry parameters are:


\begin{itemize}
    \item Logistic. Dependence parameter $\texttt{r}$ between (0,1]. Smaller \texttt{r} implies higher dependence.
    \item Asymmetrix logistic. Dependence parameter r, as in (symmetric) Logistic. Asymmetry parameters are $\texttt{t}_1$ and $\texttt{t}_2$. Indepence if any of $\texttt{t}_1, \texttt{t}_2$ are 0 or $\texttt{r = 1}$. For complete dependence $\texttt{t}_1= \texttt{t}_2$=1 and $\texttt{r} \rightarrow 0$.
    \item Husler-Reiss. Dependence parameter \texttt{r} $\in (0, \infty)$. Full dependence as $\texttt{r}$ $\rightarrow \infty$, and independence as $\texttt{r} \rightarrow 0$
    \item Negative logistic. Dependence parameter $\texttt{r}$ $>0$. Higher $\texttt{r}$ implies higher dependence. 
    \item Asymmetric negative logistic. Dependence parameter $\texttt{r}>0$ and asymmetry parameters $\texttt{t}_1,\texttt{t}_2$ $\in (0, 1]$. Indepence if any of $\texttt{t}_1,\texttt{t}_2, \texttt{r}$  approaches 0. Complete dependence if $\texttt{t}_1,\texttt{t}_2 = 1,1$ and $\texttt{r} \rightarrow \infty$.
    \item Bilogistic. Parameters $\alpha, \beta$. When $\alpha = \beta$ the model is equivalent to logistic with dependence parameter $\texttt{r} =\alpha$. As in logistic, when $\alpha = \beta =  \texttt{r} \rightarrow 0$ the model tends to complete dependence. Independence as either both tends to 1, or one is fix and other tends to 1.
    \item Negative bilogistic Parameters $\alpha, \beta$. When $\alpha = \beta$ the model is equivalent to negative bilogistic with dependence parameter $\texttt{r} =1\alpha$. When $\alpha = \beta \rightarrow 0$ the model tends to complete dependence. Independence as either both tends to $\infty$, or one is fix and other tends to $\infty$.
    \item Coles-Tawn. Parameters $\alpha, \beta > (0,0)$. As $\alpha = \beta \rightarrow \infty$ the model shows complete dependence. Independence as either both tends to 0, or one is fix and other tends to 0.
    \item Asymmetric mixed distribution. Parameters $\alpha, \beta$ fulfill the following conditions: $\alpha$ and $\alpha + 3 \beta > 0$, and $\alpha + 2\beta$, $\alpha + \beta \leq 1$ As $\beta$ is fix, the strength of dependence increases with $\alpha$. Complete dependence is not achievable. Independence as $\alpha = \beta = 0$.
  \end{itemize}
```{r 2.2}
par(mfrow=c(2,2))
sim1 = rbvevd(200,dep=0.001, model = "hr") #HR
plot(sim1)
sim2 = rbvevd(200,dep=5, model = "hr") #HR
plot(sim2)

sim3 = rbvevd(200,dep=0.001, asy=c(0.5,1), model = "aneglog") #HR
plot(sim3)
sim4 = rbvevd(200,dep=5,asy=c(0.5,1), model = "aneglog") #HR
plot(sim4)
#Hustler Reiss
(cor(sim1[,1],sim1[,2]))
(cor(sim2[,1],sim2[,2]))

(cor(sim1[,1],sim1[,2], method = "kendall"))
(cor(sim2[,1],sim2[,2], method = "kendall"))

(cor(sim1[,1],sim1[,2], method = "spearman"))
(cor(sim2[,1],sim2[,2], method = "spearman"))
# "A neglog"
(cor(sim3[,1],sim3[,2]))
(cor(sim4[,1],sim4[,2]))

(cor(sim3[,1],sim3[,2], method = "kendall"))
(cor(sim4[,1],sim4[,2], method = "kendall"))

(cor(sim3[,1],sim3[,2], method = "spearman"))
(cor(sim4[,1],sim4[,2], method = "spearman"))

```
\newpage
For the bivariate cases we have:
\begin{itemize}
    \item Bivariate Extreme Value Logistic Copula: dependence parameter $r \geq 1$. 
    \item Bivariate Extreme Value Asymmetric Logistic Copula dependence parameter $r \leq 1$, asymmetry parameters $\theta, \phi \in [0,1]$. If both are $1$ then this is a symmetric logistic.
    \item Bivariate Extreme Value Mixed Model Copula: parameter $0\leq\theta\leq1$
    \item Bivariate Extreme Value Asymmetric Mixed Model Copula: parameters $\theta, \phi$. These fulfill the following conditions, $\theta \geq 0$, $\theta + 3\phi \geq 0$, $\theta+\phi \leq 1$, $\theta + 2\phi \leq 1$.
    \item Bivariate Extreme Value Spline Copula. Parameter: a spline function for the dependence function. 
\end{itemize}
As for the multivariate copulas we have
\begin{itemize}
    \item Multivariate Extreme Value Asymmetric Logistic Copula: $\mathbf{\theta}$, matrix with asymmetry parameters. $\mathbf{r}$, vector with dependency parameters. Terms and conditions apply. 
    \item Multivariate Extreme Value Gumbel Copula. Dependence parameter $r \geq 1$.
    \item Multivariate Clayton Copula. Dependence parameter $\theta > 0$ 
    \item Multivariate Frank Copula. Parameter $\alpha > 0$
\end{itemize}
\subsection{Exercise 2.4}
```{r Exercise 2.4}
cop1 = NewBEVAsyLogisticCopula(2, 1, 0.5)
approxcop1 = GetApprox(cop1)
RV1 = GenerateRV(approxcop1, 200)
plot(RV1)

cop1 = NewBEVAsyLogisticCopula(10, 2, 1)
approxcop1 = GetApprox(cop1)
RV1 = GenerateRV(approxcop1, 200)
plot(RV1)

```
